Large-scale harmonic patterns clustering

### 2010-03-05 Paper/Paper.tex
improved, new table and fig

Working with large data amount of audio is a goal and a challenge at the
same time. One of the most common task is to find the nearest neighbor,
based on some distance, to a given segment of music. This can be viewed
as \textit{retrieval} as you are looking for some specific target
based on some query. $k$-nearest neighbors ($k$nn) algorithms have been
designed for that purpose, and recent improvements let them scale to
considerable data set sizes \cite{Datar2004}. The \textit{Shingles}
were introduced in that framework \cite{Casey2006}. A \textit{Shingle}
is a pattern of stacked chroma features. Similarity between
\textit{Shingles} coresspond to comparing harmonic patterns. For this
reason, they were used for cover or remix recognition \cite{Casey2007}.

That being said, nearest neighbor search is only one aspect of the
analysis of large-scale music data sets. Finding $k$ neighbors for a
particular query does not tell you much about the music space.
Clustering is a larger and more flexible framework where the goal is
to divide the space in meaningful blocks. The starting point is not
a query, but all the data set itself. In this work, we bring the
\textit{Shingle} idea into that framework in order to discover
meaningful and typical music harmonic patterns. The underlying assumption
is that the music space is articulated around a certain number of
frequent patterns.

Note that clustering and neighbor search is not totally unrelated.
In practise, we would hope that close neighbors live in one or two
close clusters, thus clustering can be a preprocess step. Also,
knowing all $k$ neighbors of each sample let you do graph clustering,
a subset of the clustering algorithms.

@@

@@

\begin{algorithm}
\caption{Pseudocode of Vector Quantization}
\label{algo:vq}
\begin{algorithmic}
\STATE$l$ learning rate
\STATE$\{P_n\}$ set of patches
\STATE$\{C_k\}$ codebook of $K$ codes
%\STATE $m \leftarrow b$
\REQUIRE $0 < l \leq 1$
\FOR{$nIters$}
\FOR{$p \in \{P_n\}$}
\STATE$c \leftarrow min_{c \in C_k} dist(p,c)$
\STATE$c \leftarrow c + (p - c) * l$
\ENDFOR
\ENDFOR
\RETURN $\{C_k\}$
\end{algorithmic}
\end{algorithm}

@@

\subsection{Data Size}
We are interesting to see how the amount of training data changes
the quality of our code.
We experiment with different data sizes for a fixed size of codebook and a 
fixed code length. The data points are chose randomly from the whole
data set.
Code length: 1 bar resized to size 4. Code book size: 100
We measure the distortion on our test set. We hope for some convergence.
See results in Table \ref{tab:datasize}.

Our learning can overfit, which explain the bad scores of small datasets.
The larger the dataset, the smallest the chance to overfit.
We ar elooking for a size of dataset that guarantees us a good learning
process, while being unnecessary large.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
$10$K & $50$K & $100$K & $500$K & $2$M  \\ \hline
$0.143$ & $$ & $$ & $$ & $$\\ \hline
\end{tabular}
\end{center}
\caption{{Distortion for a code book of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $10$ thousands to $2$ millions. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.}}
\label{tab:datasize}
\end{table}
% RESULTS:    avg.       mean dist
% 10K         2.012      0.143
% 50K
% 100K
% 500K
% 2M

@@

@@

\small
\section{Acknowledgements}
Thierry is NSERC graduate fellow, or some title like that.

>>>>

Working with large amount of audio data is a goal and a challenge at the
same time. One of the most common task is to find the nearest neighbor,
based on some distance, to a given segment of music. This can be viewed
as \textit{retrieval} as you are looking for some specific target
based on some query. $k$-nearest neighbors ($k$nn) algorithms have been
designed for that purpose, and recent improvements let them scale to
considerable data set sizes \cite{Datar2004}. The \textit{Shingles}
were introduced in that framework \cite{Casey2006}. A \textit{Shingle}
is a pattern of stacked chroma features. Similarity between
\textit{Shingles} coresponds to comparing harmonic patterns. For this
reason, they were used for cover or remix recognition \cite{Casey2007}.

That being said, nearest neighbor search is only one aspect of the
analysis of large-scale music data sets. Finding $k$ neighbors for a
particular query does not tell you much about the music space.
Clustering is a larger and more flexible framework where the goal is
to divide the space in meaningful blocks. The starting point is not
a query, but all the data itself. In this work, we bring the
\textit{Shingle} idea into that framework in order to discover
meaningful and typical music harmonic patterns. The underlying assumption
is that the music space is articulated around a certain number of
frequent patterns.

Note that clustering and neighbor search are not unrelated.
In practise, we would hope that close neighbors live in one or two
close clusters, thus clustering can be a preprocess step. Also,
knowing all $k$ neighbors of each sample let you do graph clustering,
a subset of the clustering algorithms.

A major component of this work is scalability. We experiment with a training
set of $43$ thousands song, but everything scales linear in time
relative to the number of samples. We could imagine, as an extension
of this work, an algorithm ``listening'' to all music it can find online.
Time would be an issue, but not memory.

@@

\subsection{Definitions}
A pattern, or a patch, is a \textit{Shingle} taken from an actual song.
A code is similar to a pattern, but is the result of the clustering
algorithm and most likely does not come from an actual song.
Distortion is our measure of distance between a pattern and a code.

@@

\begin{algorithm}
%\caption{Pseudocode of Vector Quantization}
\label{algo:vq}
\begin{algorithmic}
\STATE$l$ learning rate
\STATE$\{P_n\}$ set of patches
\STATE$\{C_k\}$ codebook of $K$ codes
%\STATE $m \leftarrow b$
\REQUIRE $0 < l \leq 1$
\FOR{$nIters$}
\FOR{$p \in \{P_n\}$}
\STATE$c \leftarrow min_{c \in C_k} dist(p,c)$
\STATE$c \leftarrow c + (p - c) * l$
\ENDFOR
\ENDFOR
\RETURN $\{C_k\}$
\caption{{Pseudocode of Online Vector Quantization. Note that we can replace
the number of iteration by a threshold on the distortion over some test set.}}
\end{algorithmic}
\end{algorithm}

@@

\subsection{Data Size}
We are interesting to see how the amount of training data changes
the quality of our code.
We experiment with different data sizes for a fixed size of codebook and a 
fixed code length. The data points are chose randomly from the whole
data set.
Code length: 1 bar resized to size 4. Code book size: 100
We measure the distortion on our test set. We hope for some convergence.
See results in Figure \ref{fig:sizes}.

Our learning can overfit, which explain the bad scores of small datasets.
The larger the dataset, the smallest the chance to overfit.
We are looking for a size of dataset that guarantees us a good learning
process, while being unnecessary large.

$500$ samples per code in the code book seems a good data size, see
Figure \ref{fig:sizes} for size $50$K.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{data_sizes}
\end{center}
\caption{{Distortion for a code book of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $10$ thousands to $2$ millions. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.}}
\label{fig:sizes}
\end{figure}

\iffalse
\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
size & $5$K & $10$K & $50$K & $100$K & $500$K & $2$M  \\ \hline
distortion & $$ & $0.143$ & $0.142$ & $0.142$ & $0.142$ & $$\\ \hline
\end{tabular}
\end{center}
\caption{{Distortion for a code book of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $10$ thousands to $2$ millions. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.}}
\label{tab:datasize}
\end{table}
\fi
% RESULTS:    
% 1bar s4     avg.       mean dist
% 1K          2.30777    0.150370
% 5K          1.88822    0.145327
% 10K         2.00349    0.143567
% 50K         2.59714    0.142257
% 100K        1.89185    0.142195
% 250K        2.07822    0.141796
% 500K        1.88141    0.142081
% 1M          1.91990    0.142118
% 2M          2.01141    0.141596

% 1bar s4 first X samples
% 1K          2.05621    0.157142
% 5K          2.01956    0.150045
% 10K         1.95797    0.148357
% 50K         1.91257    0.146069
% 100K        1.98744    0.145765
% 250K        1.95362    0.145819
% 500K        2.00899    0.145772
% 1M          2.09856    0.145924
% 2M          1.91375    0.146065

% 0bar s4 (and test set same, we could do 1bar s4 in test set...)
% 1K          2.10588    0.155474
% 5K          1.92358    0.149619
% 10K         2.00653    0.148056
% 50K         1.89443    0.146328
% 100K        1.92187    0.146004
% 250K        1.93067    0.146024
% 500K        1.91913    0.145962
% 1M          2.24550    0.145873
% 2M          2.01653    0.145955

% 2bar s8     
% 1K          3.27198    0.171300
% 5K          3.56894    0.165960
% 10K         2.61057    0.165063
% 50K         3.03945    0.162439
% 100K        2.89310    0.161588
% 250K        2.90660    0.161910
% 500K        2.69134    0.161233
% 1M          3.05344    0.161057
% 2M          2.74493    0.161068

@@

See code book size experiments in Table \ref{tab:cbsize}.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
\# codes & \# samples & avg. dist \\ \hline \hline
1 & $500$ & \\
10 & $5$K &\\
50 & $25$K &\\
100 & $50$K &\\
200 & $100$K & \\
500 & $250$K & \\
1000 & $500$K & \\
5000 & $2.5$M & \\ \hline
\end{tabular}
\end{center}
\caption{{Code book size.}}
\label{tab:cbsize}
\end{table}

@@

\small
%\section{Acknowledgements}
%Thierry is NSERC graduate fellow, or some title like that.
